---
title: "group by and summarize"
output: github_document
---

```{r}
library(tidyverse)
library(knitr)

knitr::opts_chunk$set(
  fig.width = 6, 
  fig.asp = .6, 
  out.width = "90%"
)
```

Load in a dataset that we'll use often

```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```

lubridate package creates a month variable - rounds date down to the first day of the month 

## `group_by`

```{r}
weather_df %>% 
  group_by(name)

weather_df %>% 
  group_by(name, month)
```

Lists the number of groups I have at the top (nothing much else changes)

## count some things

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(n_obs = n())

weather_df %>% 
  group_by(name, month) %>% 
  summarize(n_obs = n())
```

Column for the group that I defined and then column two shows observations

Then column for each part of the group and then n obs

It's nice that summarize produces a data frame

`table` doesn't and that's bad (can't plug this into ggplot)

```{r}
weather_df %>% 
  pull(name) %>% 
  table()

weather_df %>% 
  janitor::tabyl(month, name)
```

This is the same information as group by then summarize, but the first one is a data frame. 

This is more human readable for NOW but the other option is more helpful to then pipe into other things

## More general summaries 

Let's count, take means, look at SDs ...

```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(
    n_obs = n(), 
    mean_tmax = mean(tmax)
  )

weather_df %>% 
  group_by(name) %>% 
  summarize(
    n_obs = n(), 
    mean_tmax = mean(tmax, na.rm = TRUE)
  )

weather_df %>% 
  group_by(name) %>% 
  summarize(
    n_obs = n(), 
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmax = median(tmax, na.rm = TRUE), 
    sd_tmax = sd(tmax, na.rm = TRUE)
  )

weather_df %>% 
  group_by(month, name) %>% 
  summarize(
    n_obs = n(), 
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmax = median(tmax, na.rm = TRUE), 
    sd_tmax = sd(tmax, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
  geom_point() + 
  geom_line()

weather_df %>% 
  group_by(name) %>% 
  summarize(
    n_obs = n(), 
    mean_tmax = mean(tmax, na.rm = TRUE)) %>% 
    knitr::kable(digits = 2)
```

Waikiki - NA for tmax (because of a missing observation in the original df)

## grouped `mutate`

```{r}
weather_df %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE)
  )
```

Added a column that calculates the mean of tmax across all observations 

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE)
  )
```
group --> mutate, adds column to dataframe that is very group-specific 

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE), 
    centered_tmax = tmax - mean_tmax
  )
```

subtract group-specific mean from the intividual tmax 

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE), 
    centered_tmax = tmax - mean_tmax
  ) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
  geom_point()
```

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    tmax_rank = min_rank(tmax)
  )
```

ranked coldest days in Central Park, in x name 

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    tmax_rank = min_rank(tmax)
  ) %>% 
  filter(tmax_rank < 2)

weather_df %>% 
  group_by(name) %>% 
  mutate(
    tmax_rank = min_rank(desc(tmax))
  ) %>% 
  filter(tmax_rank < 2)
```

Variety of different ranking functions; deal with ties in different ways, use min rank and if ties really matter you can see help files 

Windowed functions depend on other things within the df (vs. just absolute value)

Lagged variables (lagged effect of some exposure down the road)

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    lagged_tmax = lag(tmax, n=1)
  )
```

NA value #1 because there are no dates before 
You need the grouping structure in place or else the first day from Waikiki would pull from lag temp max from Waterhole which is super inaccurate 

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    lagged_tmax = lag(tmax, n=1), 
    tmax_diff = tmax - lagged_tmax) %>% 
  summarize(diff_sd = sd(tmax_diff, na.rm = TRUE))
```

Day-to-day fluctuations --> 4 degrees in Central Park! 

## Limitations 

What if my "summary" is a linear model ...

```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(cor_tmin_tmax = cor(tmin, tmax, use = "complete"))

weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  lm(tmax ~ tmin, data = .)


```

This is not a single number summary 
- Group by and summarize work well with single number summaries 


lm = linear model 

weather_df %>% 
  group_by(name) %>% 
  summarize(lm = lm(tmax ~ tmin) 

THIS DOESN'T WORK 





