---
title: "ggplot 1"
output: github_document
---

Load in a dataset that we'll use often

```{r}
library(tidyverse)
library(ggridges)
```


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Scatterplot

tmax vs. tmin 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

you can save ggplots 

```{r}
ggp_tmax_tmin = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_tmax_tmin
```

## Let's fancy it up 

Add... color? lines? other stuff? 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() 

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE)

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE)

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .3) +
  geom_smooth(se = FALSE)

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

```

By moving color = name from global to the geometry, you put color only in the points, but the line is no longer applying to the individual lines, but to the global piece

Let's make one more scatterplot

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, size = prcp)) +
  geom_point(alpha = 0.3)+
  facet_grid(. ~ name) +
  geom_smooth(se = FALSE)

weather_df %>% 
  ggplot(aes(x = date, y = tmax)) +
  geom_point(aes(size = prcp), alpha = 0.3)+
  facet_grid(. ~ name) +
  geom_smooth(se = FALSE)

weather_df %>% 
  ggplot(aes(x = date, y = tmax)) +
  geom_point(aes(size = prcp), alpha = 0.3)+
  facet_grid(. ~ name) +
  geom_smooth()
```

se = FALSE (removing standard error bars across the smooth line)

## Use data manipulation as part of this 

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax = tmax * (9/5)+ 32, 
    tmin = tmin * (9/5) +32
  ) %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

## Stacking geoms 

Which geoms do you want? 

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_smooth()
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex()

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_bin_2d()
```

Changes data instensity (how many datapoints at each place); how dense data is at any particulat point

## Univariate plots 

```{r}
weather_df %>% 
  ggplot(aes(x = tmax)) +
  geom_histogram()

weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram()

weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() + 
  facet_grid(. ~ name)
```

Other plots 

```{r}
weather_df %>% 
  ggplot(aes(x = tmax)) + 
  geom_density()

weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) + 
  geom_density()

weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) + 
  geom_density(alpha = .3)
```

Still with tmax and name

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot()
```

Some people like violin plots? 

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin()
```

What about ridges

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) + 
  geom_density_ridges()

weather_df %>% 
  ggplot(aes(x = tmax, y = name)) + 
  geom_density_ridges(alpha = .8)

weather_df %>% 
  ggplot(aes(x = tmax, y = name)) + 
  geom_density_ridges(alpha = .8, scale = .8)
```




