---
title: "ggplot 2"
output: github_document
---

Load in a dataset that we'll use often

```{r}
library(tidyverse)
library(viridis)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6, 
  fig.asp = .6, 
  out.width = "90%"
)
```


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Start with a familiar one

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three stations", 
    x = "Minimum daily temp (C)", 
    y = "Maximum daily temp (C)", 
    caption = "Data from moaa package with three stations"
  )
```

## Scales 

Change how you map variables onto your graphic 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three stations", 
    x = "Minimum daily temp (C)", 
    y = "Maximum daily temp (C)", 
    caption = "Data from moaa package with three stations"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15 C", "0", "15")
  ) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right"
    )

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three stations", 
    x = "Minimum daily temp (C)", 
    y = "Maximum daily temp (C)", 
    caption = "Data from moaa package with three stations"
  ) +
  scale_color_hue(
    name = "Location", 
    h = c(100, 300)) +
    scale_color_viridis_d()
```

Scale, on this piece (x, y, or color), and its a continuous var 

## Themes

Plot elements that aren't directly tied to data, geometry, scales

Things that make your plot look nicer 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three stations", 
    x = "Minimum daily temp (C)", 
    y = "Maximum daily temp (C)", 
    caption = "Data from moaa package with three stations"
  ) +
    scale_color_viridis_d() + 
  theme_minimal() +
  theme(legend.position = "bottom") 
```

It cares what order you do certain steps in; the bw command is a delete the prior theme and do this instead; so that "trumps" legend position if you do that second

Other themes: theme_minimal()
theme_bw()
ggthemes::theme_excel() --> looks like it came out of Excel 

## `data` in geoms

```{r}

central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

waikiki %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line()

waikiki %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```









