---
title: "Simple document"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)

```

## Import some data

I want to import `FAS_litters.csv`. 

```{r}
litters_df = read_csv("data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df) 

pups_df = read_csv("data/FAS_pups.csv")
pups_df = janitor::clean_names(pups_df)
```

## `select`

Let's select some columns 

```{r}
select(litters_df, group, litter_number) 
select(litters_df, group, gd0_weight, gd18_weight)

select(litters_df, group, gd0_weight:gd_of_birth)

select(litters_df, -litter_number)
```

can also do starts_with, ends_with, or contains to select certain variables 

Can also rename variables with this: 

```{r}
select(litters_df, GROUP = group, litter_number)

rename(litters_df, GROUP = group)

select(litters_df, litter_number, everything())
```

listing litter_number first followed by everything () lets you move variables earlier/ later 

relocate (litters_df, litter_number) <-- does the same thing

Test exercise: 

```{r}

pups_df 

select(pups_df, litter_number, sex, pd_ears)

```

If you selet only 1 column, it spits out a data frame with only one column 
Does the same thing as pull

```{r}
select(litters_df, group)

pull(litters_df, group)
```

## `filter`

Let's get rid of rows

```{r}
filter(litters_df, gd_of_birth == 20)
filter(litters_df, group == "Con7")

filter(litters_df, gd0_weight < 23)

filter(litters_df, pups_survive != 4)
filter(litters_df, pups_survive == 4)

filter(litters_df, !(group == "Con7"))

filter(litters_df, group %in% c("Con7", "Con8"))

filter(litters_df, group == "Con7", gd_of_birth == 20)
```

Two equal signs is a logical operator vs. one = assigns a variable 

! before something is "not" 

%in% is either this or this (is in this set of things in the c)

& can also be , 

| is or filter(litters_df, group == "Con7" | gd_of_birth == 20)

Filter out NA/ missing values --> drop_na(litters_df) - drops all rows with any missing values OR drop_na(litters_df, gd_of_birth) - drops all rows with missing values for this variable 

## `mutate` 

Let's add or change columns! 

```{r}
mutate(litters_df, weight_change = gd18_weight - gd0_weight)
```

Use litters data frame and use this new variable 

Can often separate across lines 

```{r}
mutate(
  litters_df, 
  weight_change = gd18_weight - gd0_weight, 
  group = str_to_lower(group))
```

str_to_lower - makes all values in that variable lowercase 

## `arrange` 

Let's arrange the data. 

```{r}
arrange(litters_df, gd0_weight)
arrange(litters_df, desc(gd0_weight))

arrange(litters_df, gd_of_birth, gd0_weight) 
```

Organizes first by gd of birth and then gd0 weight 

## Pipes

```{r}
litters_data_raw = read_csv("data/FAS_litters.csv")
litters_clean_names = janitor::clean_names(litters_data_raw)
litters_select = select(litters_clean_names, group, pups_survive)
litters_filtered = filter(litters_select, group == "Con7")

litters_filtered

```

TRY AGAIN with Pipes

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv") %>%
  janitor:: clean_names() %>%
  select(group, pups_survive) %>%
  filter(group == "Con7")

litters_df

litters_try3 = 
  read_csv("data/FAS_litters.csv") %>%
  janitor:: clean_names() %>%
  select(-pups_survive) %>%
mutate(
    weight_change = gd18_weight - gd0_weight, 
    group = str_to_lower(group) 
) %>%
  drop_na(weight_change) %>% 
  filter(group %in% c("con7", "con8")) %>%
  select(litter_number, group, weight_change, everything())
```
  
You don't have to repeat the data frame - that's built in 


Question - where does the dataframe go? if you use litters_df, is that edited? 



