---
title: "ggplot 2"
output: github_document
---

Load in a dataset that we'll use often

```{r}
library(tidyverse)

library(patchwork)
library(viridis)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6, 
  fig.asp = .6, 
  out.width = "90%"
)
```


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Start with a familiar one

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three stations", 
    x = "Minimum daily temp (C)", 
    y = "Maximum daily temp (C)", 
    caption = "Data from moaa package with three stations"
  )
```

## Scales 

Change how you map variables onto your graphic 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three stations", 
    x = "Minimum daily temp (C)", 
    y = "Maximum daily temp (C)", 
    caption = "Data from moaa package with three stations"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15 C", "0", "15")
  ) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right"
    )

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three stations", 
    x = "Minimum daily temp (C)", 
    y = "Maximum daily temp (C)", 
    caption = "Data from moaa package with three stations"
  ) +
  scale_color_hue(
    name = "Location", 
    h = c(100, 300)) +
    scale_color_viridis_d()
```

Scale, on this piece (x, y, or color), and its a continuous var 

## Themes

Plot elements that aren't directly tied to data, geometry, scales

Things that make your plot look nicer 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three stations", 
    x = "Minimum daily temp (C)", 
    y = "Maximum daily temp (C)", 
    caption = "Data from moaa package with three stations"
  ) +
    scale_color_viridis_d() + 
  theme_minimal() +
  theme(legend.position = "bottom") 
```

It cares what order you do certain steps in; the bw command is a delete the prior theme and do this instead; so that "trumps" legend position if you do that second

Other themes: theme_minimal()
theme_bw()
ggthemes::theme_excel() --> looks like it came out of Excel 

## `data` in geoms

```{r}

central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

waikiki %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line()

waikiki %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```

## `patchwork`

```{r}
ggp_tmax_tmin = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  theme(legend.position = "none") 

ggp_prcp_dens = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .3) +
  theme(legend.position = "none") 

ggp_tmax_date = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_smooth() + 
  theme(legend.position = "bottom")

(ggp_tmax_tmin + ggp_prcp_dens) / ggp_tmax_date

```

## data manipulation 

quick example on factors

```{r}
weather_df %>% 
  mutate(
    name = fct_reorder(name, tmax)
  ) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot()
```

R assumes: 
categorical variable --> factor variable --> alphabetical order

What about tmax and tmin 

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin, 
    names_to = "obs", 
    values_to = "temperature") %>% 
  ggplot(aes(x = temperature, fill = obs)) + 
  geom_density(alpha = .3) + 
  facet_grid(. ~ name)
```

```{r}
pulse_df = 
  haven::read_sas("data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m, 
    names_to = "visit", 
    values_to = "bdi", 
    names_prefix = "bdi_score_"
  ) %>% 
  mutate(visit = recode(visit, "bl" = "00m"))

pulse_df %>% 
  ggplot(aes(x = visit, y = bdi)) + 
  geom_boxplot()

pulse_df %>% 
  ggplot(aes(x = visit, y = bdi)) + 
  geom_point() + 
  geom_line(aes(group = id))
```



